\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\linespread{1.2}
\usepackage{babel}
\usepackage{amsthm, bm}
\usepackage{graphicx}
\usepackage{cancel}
\usepackage{comment}
\usepackage{enumerate}
\usepackage{booktabs}
\usepackage{float}
\usepackage{subcaption}
\usepackage{geometry}
\geometry{a4paper,top=2.5cm,bottom=2.5cm,left=2cm,right=2cm}
\usepackage{caption}
\usepackage[version=4]{mhchem}
\usepackage{stmaryrd}
\usepackage{hyperref}
\usepackage{float}
\hypersetup{colorlinks=true, linkcolor=blue, filecolor=magenta, urlcolor=cyan,}
\urlstyle{same}
\usepackage{bbold}

\title{Timeserie Econometrics 3 - Problem Set 1}
\author{Alex Basov, Lorenzo Gorini}
\date{\today}

\begin{document}
	
	\maketitle
	\vspace{1cm}
	
	\section*{Lab Exercises}
	\subsection*{Question 1: VAR and impulse response functions (IRF)}
		Refer to the 3 code files for the solution. Particularly, the 2 python files are for data cleaning and merging. The R file "model.R" is to reply to the following questions.
		Here we only report the main results.
		
		\subsubsection*{Part 1}
		\textit{Before fitting a trivariate VAR(p) for the 3-variable you just collect, briefly discuss if there are any transformations needed for the raw data}
		
		\subsubsection*{Part 2}
		\textit{Estimate the model based on transformed series, show the estimates of the coefficients and of the covariance matrix you obtain}
		
		\subsubsection*{Part 3}
		\textit{Use the estimates above as DGP to design a Monte Carlo experiment. Assess the accuracy of the AIC and BIC criterion in selecting the lag order of a VAR. Comment on the results you obtain.}
		
		\subsubsection*{Part 4}
		\textit{Suppose now you want to identify and analyze the effects of a monetary policy shock:
			\begin{itemize}
				\item Comment on the ordering of the variables you would use to recursively identify the shock.
				\item Obtain IRFs estimates from both VAR and Local Projections method (Jorda (2005,AER)). Comment on the differences you find
			\end{itemize}
		}
\end{document}
